Re-run, Repeat, Reproduce, Reuse, Replicate:\\Transforming Code into Scientific Contributions

### 2017-07-24 draft.tex
Modified slightly the conclusion

But precisely because it is easy to believe that if a program runs once and gives the expected
results it will do so forever, crucial steps to transform working code into
meaningful scientific contributions are rarely undertaken
\citep{Sandve:2013,Schwab:2000}. Computer science is plagued by replication
problems, in part, precisely because it seems impervious to them. A
program can fail as a scientific contribution in many different ways for many
different reasons. Borrowing the terms coined by \citeauthor{Goble:2016}
\citep{Goble:2016}, for your program to contribute to science, you have to make
sure it is re-runnable (R$^1$), repeatable (R$^2$), reproducible (R$^3$),
reusable (R$^4$) and replicable (R$^5$).\\

Let us illustrate this with a small example, a random walk written in Python:\\

@@

def walk():
    promenade, total = [], 0
    for i in range(10):
        step = +1 if random.uniform(-1,+1) > 0 else -1
        total += step
        promenade.append(total)
    return promenade

@@

%FIXME: can still be improved
Through a small random walk example using Python, we've seen a variety of pitfalls that are easy to remain unaware of while writing scientific code. The R$^0$ form, while easy to write and simple to understand is not a good scientific code while the full-fledge R$^4$ form is much preferred but requires a fair amount of work that might be too much of a burden. R$^3$ form seems to be an acceptable compromise and should be accepted as the minimum scientific standard \citep{Wilson:2017}. This means this should be actually checked by reviewers and publishers when code is part of a work worth to be published. But it's hardly the case today.\\

>>>>

But precisely because it is easy to believe that if a program runs once and
gives the expected results it will do so forever, crucial steps to transform
working code into meaningful scientific contributions are rarely undertaken
\citep{Sandve:2013,Schwab:2000}. Computer science is plagued by replication
problems, in part, precisely because it seems impervious to them. A program can
fail as a scientific contribution in many different ways for many different
reasons. Borrowing the terms coined by \citeauthor{Goble:2016}
\citep{Goble:2016}, for your program to contribute to science, you have to make
sure it is re-runnable (R$^1$), repeatable (R$^2$), reproducible (R$^3$),
reusable (R$^4$) and replicable (R$^5$). Let us illustrate this with a small
example, a random walk written in Python:\\

@@

def walk():
ea    promenade, total = [], 0
    for i in range(10):
        step = +1 if random.uniform(-1,+1) > 0 else -1
        total += step
        promenade.append(total)
    return promenade

@@

%FIXME: can still be improved
Through a small random walk example using Python, we've seen a variety of
pitfalls that are easy to remain unaware of while writing scientific code. The
R$^0$ form, while easy to write and simple to understand is not a good
scientific code. Adding a single comment, we transformed this code into the
R$^1$ version (rerunnable) that partially solved the very common dependencies
problem and will allow us to re-run the program two months or two years from
now. In the meantime, the R$^2$ version (repeatable) guarantee us (to some
extent) we can faithfully repeat our experiment from one run to the other.  Of
course, the full-fledge R$^4$ (reusable) form is much preferred but it requires
a fair amount of work that might be too much of a burden. R$^3$ (reproducible)
form seems to be an acceptable compromise and should be accepted as the minimum
scientific standard \citep{Wilson:2017}. This means this should be actually
checked by reviewers and publishers when code is part of a work worth to be
published. But it's hardly the case today.\\
